\documentclass{article}
\usepackage{iclr2019_conference,times}
\iclrfinalcopy

\input{math_commands.tex}

\usepackage{amsthm}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
%\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amsthm}

\newcommand{\kernel}{winning ticket}
\newcommand{\kernels}{winning tickets}
\newcommand{\commentt}[1]{}

\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{amsmath}

\newcommand\newsubcap[1]{\phantomcaption%
       \caption*{\figurename~\thefigure(\thesubfigure): #1}}

% Packages.

\definecolor{shadeOne}{rgb}{0.9,0.95,0.95}
\definecolor{shadeTwo}{rgb}{0.99,0.99,0.99}

\usepackage[bordercolor=shadeOne, backgroundcolor=shadeTwo, linecolor=red, textwidth=0.45in, textsize=tiny]{todonotes}

\input{commenting}

\title{The Lottery Ticket Hypothesis:\\Finding Sparse, Trainable Neural Networks}


\author{Jonathan Frankle\\MIT CSAIL\\\texttt{jfrankle@csail.mit.edu} \And Michael Carbin\\MIT CSAIL\\\texttt{mcarbin@csail.mit.edu}}


\begin{document}

\maketitle

\begin{abstract}
Neural network pruning techniques can reduce the parameter counts of trained networks by over 90\%, decreasing storage
requirements and improving computational performance of inference without compromising accuracy.
However, contemporary experience is that the sparse architectures produced by pruning are difficult to train
from the start, which would similarly improve training performance.

\vspace{.5em}
We find that a standard pruning technique
naturally uncovers subnetworks whose initializations made them capable of training effectively.
Based on these results, we articulate
the \emph{lottery ticket hypothesis}: dense, randomly-initialized, feed-forward networks contain
subnetworks (\emph{winning tickets}) that---when trained in isolation---reach test accuracy comparable to the original network in a similar number of iterations.
The winning tickets we find
have won the initialization lottery: their connections have initial weights that make training particularly effective.

\vspace{.5em}
We present an algorithm to identify winning tickets and a series of experiments that support
the lottery ticket hypothesis and the importance of these fortuitous initializations. 
We consistently find winning tickets that are less than 10-20\% of the size of
several fully-connected and convolutional feed-forward architectures for MNIST and CIFAR10.
Above this size, the winning tickets that we find learn faster
than the original network and reach higher test accuracy.

\end{abstract}


\input{intro}

\input{mnist-oneshot}

\input{convolutional}


\input{dropout}

\input{larger}

\input{discussion}

\input{limitations}


\input{related}


\bibliography{local}
\bibliographystyle{iclr2019_conference}

\newpage

\begin{appendix}


\input{acknowledgements}

\input{app-han}

\input{app-early-stopping}

\input{app-train}

\input{app-random-sparsity}

\input{examining}

\input{mnist-appendix}

\input{convolutional-appendix}

\input{appendix-resnet}


\end{appendix}

\end{document}
